Design and Implementation of Probabilistic Programming Language Anglican

### 2017-02-20 paper/paper.tex
update

\hy{Need to be revised.}

Programs in Anglican define probability distributions 
over sequences of sampled values through their executions.

Imagine an interpreter of Anglican that runs the deterministic parts of a Anglican program 
according to its standard semantics, executes its \texttt{sample} form by
generating a random sample, and treats its $\texttt{observe}$ form by skip.
In addition to this execution, the interpreter keeps a log that records
information about all the \texttt{sample} and \texttt{observe} forms encountered 
during execution. The information kept for \texttt{sample} is a triple $(F,x,\alpha)$ of
the following data:
\begin{itemize}
        \item $F$ is one of the built-in primitive probability distributions, 
                such as the standard normal, for which we have
                the probability density or probability mass function $p_F$.
        \item $x$ is a value sampled from the distribution $F$.
        \item $\alpha$ is an address that uniquely and systematically identifies the random 
                choice made; this address is exploited in inference algorithms, as
                explained later.
\end{itemize}
The information kept for \texttt{observe} is a pair $(G,y)$ where $G$
is a primitive probability distribution as $F$ above and $y$ is an observed value.
Thus, a log is just a sequence of triples $(F,x,\alpha)$ and pairs $(G,y)$.

We call the triple part of a log \textit{trace}. This part records all the random 
choices of an execution. Thus, if the trace of a log is fixed, the rest of the log 
is decided determistically by the program. A trace induces
a sequence of pairs $(G_j, y_j)$ of distributions and values of
\textit{observed} random variables. We call this sequence an \textit{image} and
denote it by $\pmb{y}$. For notational simplicity we assume that the program
always generates the same ordered set $\pmb{z}$ of \textit{results} $z_k$.

The target density $\pi(\vec{x}) := \gamma(\vec{x}) / Z$ of a program
is defined in terms of the product of the probabilities of all random
choices $\pmb{x}$ and the likelihood of all observations $\pmb{y}$
\begin{align}
  \gamma(\pmb{x}) 
  &:= 
  \prod_{i=1}^{\left|\pmb{x}\right|}
  p_{F_i}(x_i) \prod_{j=1}^{\left|\pmb{y}\right|}p_{G_j}(y_{j}).
  \label{eqn:p-trace}
\end{align}
The objective of inference in probabilistic program $\mathcal{P}$
is to discover the distribution of $\pmb{z}$.

>>>>

Anglican programs define probability distributions 
over sequences of annotated values, implicitly by means
of program execution.
Imagine an interpreter that runs the deterministic parts of a program 
according to the standard semantics, executes the \texttt{sample} form by
generating a random sample, and treats the $\texttt{observe}$ form by skip.
In addition to this execution, the interpreter also keeps a log that records
information about all the \texttt{sample} and \texttt{observe} forms encountered 
during execution. The information recorded for \texttt{sample} is 
a triple $(F,x,\alpha)$ of (i) a primitive probability distribution $F$, 
such as the standard normal, for which we have 
the probability density $p_F$; (ii) a value $x$ sampled from the distribution $F$; 
and (iii) an address $\alpha$ that uniquely and systematically identifies the random 
choice made. The information recorded for \texttt{observe} is a pair $(G,y)$ where $G$
is a primitive probability distribution as $F$ from above and $y$ is an 
observed value. Thus, a log is a sequence of triples $(F,x,\alpha)$ and 
pairs $(G,y)$.

A \textit{trace} $\pmb{x}$ is a sequence of triples $(F,x,\alpha)$. It is feasible if
it is precisely the triple part of the log of some execution. 
Such a feasible trace records all the random choices made during execution,
and uniquely determines the rest of the execution and its log, because all the non-sample
forms of Anglican programs are deterministic. In particular, each of these feasible traces
decides a sequence of pairs $(G_j, y_j)$ for observed values.
We call the sequence \textit{image of $\pmb{x}$} and denote it by~$\pmb{y}$.

A (almost surely terminating) probabilistic program defines a probability distribution 
over feasible traces $\pmb{x}$, which has the probability density 
$\pi(\pmb{x}) := \gamma(\pmb{x}) / Z$ where 
\begin{align}
  \gamma(\pmb{x}) 
  &:= 
  \prod_{i=1}^{\left|\pmb{x}\right|}
  p_{F_i}(x_i) \prod_{j=1}^{\left|\pmb{y}\right|}p_{G_j}(y_{j}).
  \label{eqn:p-trace}
\end{align}
for the image $\pmb{y}$ of $\pmb{x}$, and $Z$ is the normalization constant $Z := \int \gamma(\pmb{x})d(\pmb{x})$
of $\gamma$.\footnote{The integral and the density use the default measure obtained by the standard extension of
a $\sigma$-finite measure on finite traces where the $\sigma$-finite measure itself is defined by
a countable sum of Lebesque and counting measures.}
